Universal Anonymous Signatures

### 2022-09-19 prism-backend/research/gsac/app-related-extra.tex
Completed in-depth pass. Some changes pending.

\subsubsection{\CUASGSMDO construction.} %
Our \CUASGSMDO construction is essentially a $(\fissue^{single,F,I},\feval^0,
\finsp^{\smsg})$-\CUASGen restriction, where $\fissue^{single,F,I}$ and
$\feval^0$ are as in \figref{fig:func-restrictions}, and $\finsp^{\smsg}$, is as
described in \eref{eq:finspmsg}. Since the construction is essentially a variant
of our \CUASGen-based construction for group signatures, we describe it making
reference to \CUASGS, in \secref{ssec:related-models-gs}:

@@

>>>>

\subsubsection{\CUASGSMDO construction.} %
Our \CUASGSMDO construction is essentially a $(\fissue^{single,H,I},\feval^0,
\finsp^{\smsg})$-\CUASGen restriction, where $\fissue^{single,H,I}$ and
$\feval^0$ are as in \figref{fig:func-restrictions}, and $\finsp^{\smsg}$, is as
described in \eref{eq:finspmsg}. Since the construction is essentially a variant
of our \CUASGen-based construction for group signatures, we describe it making
reference to \CUASGS, in \secref{ssec:related-models-gs}:

@@

\subsection{Multimodal Private Signatures}
\label{ssec:related-models-mps}

Multimodal Private Signatures (or, rather, a close relative, see next) are
directly implementable as a $(\fissue^{single,H,I},\feval^{F,MPS},
\finsp^{G,MPS})$-\CUASGen restriction. Concretely, this restricts to a system
with one credential per user and no attributes in the credentials.

There are some differences, though. On the one hand, our signature anonymity
property is somehow in the middle of MPS's privacy-1 and privacy-2 variants.
Their privacy-1 variant is the traditional anonymity property of group
signatures, in which the opener is honest and the adversary can call the opening
oracle except for challenge signatures. In their privacy-2 variant, the opener
is fully corrupt (i.e., \adv~ receives \osk), but the adversary can only be
challenged with signatures that produce the same \yinsp value ($G_{j_b}(id_b)$
in MPS). In our \UAS model, we allow the adversary to corrupt openers at will
(getting ther \osk), but the challenge signatures must have been generated using
the key of an uncorrupted opener; still, \adv~can call the open oracle for
challenge signatures, as long as they have the same \yinsp value. We believe
that these differences are mostly of theoretical interest, though. In practice,
what matters is that any variant of \feval and \finsp functions ($F$ and $G_j$
in MPS) that has too much granularity in their ranges, will lead to a barely
private system. This is a known issue, for instance, in functional encryption
\cite{bsw11}.

On the other hand, anonymous issuance and the capability to customize issuance
policies are unachievable by MPS. Anonymous issuance simply does not make sense
in MPS, as it only supports one credential (without attributes) per user, and
customized issuance is achieved in \UAS via the \fissue functions, with no
equivalent in MPS -- instead, MPS can be seen as having the fixed policy of
``proving knowledge of a user secret key''. These extensions are not trivial
from a modelling point of view, and also have a big impact in the practicality
of the resulting system. Also, MPS does not include verifiable openings. While
this is easy to add, we consider it to be essential, given that the proposed
generalization of opening allows the opener to output ``any value''. Hence, the
need to ensure that the produced value is the legitimate one seems to be more
necessary than ever. Finally, our \feval function is a generalization of $F$ in
MPS, as it allows two outputs: one will be returned in the clear along with the
signature, and the other will be encrypted. Indeed, making part of the output
unencrypted allows implementing very useful variants such as anonymous
credentials with selective disclosure, as shown before.

For the model of MPS schemes, we refer to the original paper \cite{ngsy22}.

\subsubsection{\CUASMPS construction.} %
As mentioned, a $(\fissue^{single,H,I},\feval^{F,MPS},
\finsp^{G,MPS})$-restriction implements an MPS scheme where
$\fissue^{single,H,I}$ is as in \figref{fig:func-restrictions}, $\feval^{F,MPS}$
equals MPS's $F$ function, and MPS's $G$ function equals $\finsp^{G,MPS}$, both
as defined in \esref{eq:funcs-mps}.

\begin{align}
  \feval^{F,MPS}(\upk,\lbrace (\cid_i,\attrs_i) \rbrace_{i\in[n]},\msg)
  \coloneqq & \nonumber \\
  & \hspace*{-10em} \pcif F(\upk,\lbrace (\cid_i,\attrs_i) \rbrace_{i\in[n]},
    \msg) = 0: \pcreturn \bot \nonumber \\
  & \hspace*{-10em} \pcelse: \pcreturn (\ast,F(\upk,\lbrace(\cid_i,\attrs_i)
    \rbrace_{i\in[n]},\msg)) \nonumber \\
  \finsp^{G,MPS}((\cdot,\yeval^1),\upk,\cid,\attrs,\cdot) \coloneqq
  & \nonumber \\
  & \hspace*{-10em} \pcreturn G(\yeval^1,\upk,\cid,\attrs)
    \label{eq:funcs-mps}
\end{align}

With these functions, the construction is direct. \CUASMPS's setup runs
\CUASGen's \Setup, \OKeyGen, and generates the issuer key pair by running
\KeyGen and \ISet (with $\fissue^{single,H,I}$). For MPS's join, \CUASMPS runs
first \KeyGen to generate the user key pair, and then proceeds as in \CUASGen.
\Sign, \Verify and \Open, are as in \CUASGen.

Finally, note that, in MPS, $G$ is further parameterized with a value $j$, which
essentially is the result of applying $F$ to the message and signer identity,
and is not leaked to the opener nor any other third party. The equivalent in
\UAS is the $\yeval^1$ value which, in \CUASGen, is encrypted with a single-use
key pair, and is an input of \finsp.

\paragraph{Security of \CUASMPS.} %
As stated, we restrict to privacy-1 of MPS (although, as
discussed, we actually achieve something slightly stronger). Unforgeability-1 in
MPS is implied by unforgeability in \UAS (where the latter also covers
unforgeability of \fissue evaluations, plus verifiable openings).
Unforgeability-2 in MPS is equivalent to non-frameability in \UAS (where the
latter also includes framing via verifiable openings). Finally note that, in
UAS, we do not explicitly model extractability, but rather make it an inherent
requirement in the unforgeability and non-frameability definitions. Thus, our
\CUASMPS construction also ensures MPS-extractability.

\begin{theorem}[Privacy-1 of \CUASMPS]
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, then \CUASMPS satisfies privacy-1
  according to \cite{ngsy22}.
\end{theorem}

\begin{proof}  
  Note that the privacy-1 game in MPS and the signature anonymity game in \UAS
  are, in essence, equivalent. Thus, given an adversary \adv~that breaks
  privacy-1 of \CUASMPS as defined for MPS, it is easy to build an adversary
  \advB breaking signature anonymity of the underlying \CUASGen, as defined in
  \UAS.

  \advB simply leverages its own oracles in order to answer oracle queries from
  \adv: $\mathcal{O}_{read},\mathcal{O}_{write},\mathcal{O}_{CU},\mathcal{O}_{HU},
  \mathcal{O}_{sig},\mathcal{O}_{open}$ in MPS correspond to \RREG, \WREG,
  \CUGEN, \HUGEN, \SIGN and \OPEN in \UAS, respectively. The output given by
  \adv~in the first stage of the MPS game is then directly used as input to
  \advB's \CHALb oracle. Finally, \advB outputs whatever \adv~outputs.

  Since the simulation is direct, it is easy to see that, whenever \adv~wins,
  so does \advB.
  % 
  \qed
\end{proof}

\begin{theorem}[Extractability of \CUASMPS]
  If the base \CUASGen construction has non-frameability according to
  \defref{def:frame-uas}, then \CUASMPS is an extractable MPS scheme.
\end{theorem}

\begin{proof}  
  The proof is direct, as the winning conditions in the extractability
  experiment in MPS are also winning conditions in the non-frameability
  experiment for \UAS. That is, an adversary \adv~against extractability of
  \CUASMPS, as defined in MPS, can be used to build an adversary \advB breaking
  non-frameability of \CUASGen, as defined in the \UAS model.

  \advB simulates \adv's game essentially as in the privacy-1 proof. Note also
  that \adv~receives the issuer's secret key. This can be obtained by \advB via
  its \ICORR oracle. Eventually, \adv~outputs a $(F,M,\Sigma)$ tuple, which
  corresponds to $(\feval^{g,MPS},\msg,\Sig)$ (with $g=F$ in MPS) in \CUASGen,
  which \advB outputs in its own game. In addition, \advB outputs as $(\yinsp,
  \pi)$ values the output of \Open over $(\Sig,\msg)$ -- which internally
  computes $\finsp^{h,MPS}$, for $h=G$ in MPS.
  % 
  Note that the $\feval^{g,MPS}$ and $\finsp^{h,MPS}$ functions in \CUASMPS
  mimic the behaviour of MPS's $F$ and $G$, respectively. Thus, if \adv wins
  because $F$ returns $0$ when evaluated on the extracted values, then \advB
  wins at line 7 of the non-frameability game, as a legitimately computed
  $\feval^{g,MSP}$ function cannot output $0$. Similarly, if \adv~wins because
  the value returned by an honestly computed open function over the adversarial
  $(F,M,\Sigma)$ tuple does not match the output of $G$ over the extracted
  values, \advB wins at line 9: note that \advB returns the \yinsp value that
  \Open outputs (when computed honestly over \adv's output), whereas
  $\finsp^{g,MSP}=G$; thus, if both values differ in \adv's game, then they also
  differ in \advB's game.
  %
  \qed
\end{proof}

\begin{theorem}[Unforgeability-1 of \CUASMPS]
  If the base \CUASGen construction is unforgeable according to
  \defref{def:forge-uas}, then \CUASMPS satisfies unforgeability-1 as defined in
  \cite{ngsy22}.
\end{theorem}

\begin{proof}
  Given an adversary \adv~against unforgeability-1 of \CUASMPS, as defined in
  the MPS model, we build \advB that breaks unforgeability of the underyling
  \CUASGen, as defined for \UAS.

  Simulation of \adv's oracles by \advB is again direct (as in previous proofs).
  Now, suppose that \adv~wins -- i.e., the identity extracted by \adv~does not
  belong to the set of corrupt users in \adv's game. By construction of \advB
  (identities in \adv's game map to a $\cid$ for \advB), $\OWNR[\cid] \notin
  \CU$ in \advB's unforgeability game either.
  % 
  However, in order for \advB to win in its own game, we need to prove also that
  the produced identity does not belong to a user in \HU either. Suppose that it
  does. Simulation extractability ensures that the credentials extracted by
  \adv~all belong to the same user (by assumption, in \HU) -- i.e., they are
  bound to its \usk. Moreover, since \adv~wins its game, the signature output by
  \adv~has not been obtained via its \SIGN oracle. But, given that the user is
  honest, its \usk has been produced uniformly at random, and has not been
  leaked to \adv~(as the issuer is also honest). Thus, the probability that
  the extracted credentials belong to an honest user is negligible and, if
  \adv~wins its game, \advB also does.
  %
  \qed
\end{proof}

\begin{theorem}[Unforgeability-2 of \CUASMPS]
  If the base \CUASGen construction has non-frameability according to
  \defref{def:frame-uas}, then \CUASMPS satisfies unforgeability-2 as
  defined in \cite{ngsy22}.
\end{theorem}

\begin{proof}
  Given an adversary \adv~against unforgeability-2 of \CUASMPS, as defined in
  the MPS model, we build \advB that breaks non-frameability of the underyling
  \CUASGen, as defined for \UAS.

  Simulation of \adv's oracles by \advB is as in the unforgeability-1 proof.
  \advB outputs whatever \adv~outputs, adding a $(\yinsp,\iproof)$ pair, honestly
  generated from \adv's output. Note that the winning condition of \adv~
  requires that the produced signature belongs to an honest user, without
  having been produced via a call to the signing oracle. By construction of
  \advB, this directly makes \advB win in line 11 of the non-frameability game.
  %
  \qed
\end{proof}

% \begin{theorem}[Privacy-1 of \CUASMPS]
%   If the base \CUASGen construction has signature anonymity according to
%   \defref{def:sign-anonymity-uas}, then \CUASMPS satisfies privacy-1
%   according to \cite{ngsy22}.
% \end{theorem}

% \begin{theorem}[Extractability of \CUASMPS]
%   If the base \CUASGen construction is unforgeable according to
%   \defref{def:forge-uas}, then it is an extractable MPS scheme.
% \end{theorem}

% \begin{theorem}[Unforgeability-1 of \CUASMPS]
%   If the base \CUASGen construction is unforgeable according to
%   \defref{def:forge-uas}, then \CUASMPS satisfies unforgeability-1 as
%   defined in \cite{ngsy22}.
% \end{theorem}

% \begin{theorem}[Unforgeability-2 of \CUASMPS]
%   If the base \CUASGen construction has non-frameability according to
%   \defref{def:frame-uas}, then \CUASMPS satisfies unforgeability-2 as
%   defined in \cite{ngsy22}.
% \end{theorem}

\subsection{Revocable Anonymous Credentials}
\label{ssec:related-models-rac}

In \cite{cks10,ckl+15}, revocable anonymous credentials are proposed. We first
give a simple model in \figref{fig:model-rac}, extending that in
\secref{ssec:related-models-ac}. Concretely, to the functionality in
\secref{ssec:related-models-ac}, we now explicitly add the $OKeyGen$ function,
which generates the opener key pair, extend $Verify$ to receive also a list $L$
of revoked credential identifiers, and we add a function (and corresponding
oracle) $cid/\bot \gets RetrieveCredId(\osk,\utrans)$ that can extract an
identifier of the credential used in a presentation transcript \utrans. For
simplicity, we do not include the equivalent to verifiable openings, but this
would be straightforward.

\begin{figure}[ht!]
  \centering
  \scalebox{0.85}{      
    \begin{minipage}[t]{0.43\textwidth}
      \procedure[linenumbering]{$\Exp^{anon-b}_{\adv,rac}(1^\secpar)$}{%
        \parm \gets \Setup(1^\secpar) \\
        (\opk,\osk) \gets \OKeyGen(\parm) \\        
        (\ipk,\isk) \gets IssKeyGen(\parm) \\        
        b^* \gets \adv^{\mathcal{O}_{rac},\oracle{LoR}}(\parm,\ipk,\opk) \\
        \pcreturn b^*
      }
    \end{minipage}
  }
  \scalebox{0.85}{      
    \begin{minipage}[t]{.57\textwidth}
      \procedure[linenumbering]{$\Exp^{forge}_{\adv,rac}(1^\secpar)$}{%
        \parm \gets \Setup(1^\secpar) \\
        (\opk,\osk) \gets \OKeyGen(\parm) \\        
        (\ipk,\isk) \gets IssKeyGen(\parm) \\
        (\cid,D,L,\st) \gets
        \adv^{\mathcal{O}_{rac}}(\parm,\ipk,\opk) \\
        \langle \cdot,(b_0,\utrans_0)\rangle
        \gets \langle \adv(\st),\Verify(\ipk,D,L)\rangle \\
        \langle \cdot,(b_1,\utrans_1)\rangle
        \gets \langle Show(\ipk,\ATTR[\cid],D,\CRED[\cid]), \\
        \hspace*{7.25em}\Verify(\ipk,D,L)\rangle \\
        \pcif \OWNR[\cid] \notin \HU \lor
        D \not\subseteq \ATTR[\cid]: \pcreturn 0 \\
        \pcif b_0 = 1 \land \nexists j~\suchthat~(\OWNR[j] \in \CU
        \land D \subseteq \ATTR[j]): \\
        \pcind \pcreturn 1 \\
        h_{\bin} \gets RetrieveCredHandle(\osk,\utrans_{\bin}) \\
        \pcreturn h_0 \in L \lor (h_1 \notin L \land b_1 = 0)
      }       
    \end{minipage}      
  }
  \caption{Security games for revocable anonymous credentials with selective
    disclosure based on \cite{fhs19}. 
    $\mathcal{O}_{rac} \gets \lbrace \oracle{HU},\oracle{CU},\oracle{Obtain},
    \oracle{Issue},\oracle{ObtIss},\oracle{Show},\oracle{ReqCredHandle}\rbrace$.
    $\oracle{ReqCredHandle}$ is
    equivalent to \OPEN, operating on transcripts \utrans instead of signatures
    but, in addition, keeps a list of revoked users (per verifier), based on the
    queries it receives. In $\Exp^{forge}_{\adv,rac}$, \adv~wins if it
    authenticates using a revoked credential or despite not having the required
    attributes, or if it prevents an honest user from authenticating using a
    non-revoked credential.}
  \label{fig:model-rac}  
\end{figure}

\subsubsection{\CUASRAC Construction.} %
Briefly, this is achieved via a $(\fissue^1,\feval^{\dattrs,H,V,L},
\finsp^{F})$-\CUASGenInt restriction, where $\fissue^1$ is as in
\figref{fig:func-restrictions}, and $\feval^{\dattrs,H,V,L}$ and $\finsp^{H}$
are ase specified in \esref{eq:funcs-rac}.

\begin{align}
  \feval^{\dattrs,H,V,L}(\upk,((\cidi,\cidu),\attrs),\cdot) \coloneqq &
    \nonumber \\
  & \hspace*{-10em} \pcif F(V,\cidi) \in L \lor F(V,\cidu) \in L: \pcreturn \bot
  \nonumber \\
  & \hspace*{-10em} \pcelse \pcreturn (\lbrace \attrs_i \rbrace_{i\in \dattrs},V) \nonumber \\
  \finsp^{F}((\cdot,\yeval^1),\cdot,((\cidi,\cidu),\attrs),\cdot)
  \coloneqq  & \nonumber \\
  & \hspace*{-10em} \pcreturn (F(\yeval^1,\cidi),F(\yeval^1,\cidu))
  \label{eq:funcs-rac}
\end{align}

Concretely, $\feval^{\dattrs,H,V,L}$ is defined via a set of attributes \dattrs
to reveal (e.g., age, etc.), a PRF $F$, a string identifying the verifier $V$
and a list $L$ of revocation handles by $V$. Note that, these ``extra
parameters'' can be hardcoded, though: $F$ can be a fixed PRF for all verifiers,
\dattrs and $V$ are also fixed (per verifier), and $L$ can be updated per
verifier with some reasonable frequency -- e.g., daily -- so that the cost of
recomputing $\feval^{\dattrs,H,V,L}$ is not high.
%
Since credentials in \CUASGenInt have two identifiers (\cidi, chosen by the
issuer; and \cidu, chosen by the user), $\feval^{\dattrs,H,V,L}$ internally
computes $F(V,\cidi)$ and $F(V,\cidu)$. If either appears in $L$, \feval aborts.
Otherwise, it sets $\yeval^0 \gets \lbrace \attrs_i \rbrace_{i\in\dattrs}$ and
$\yeval^1 \gets V$. $\yeval^1$ is
then used by $\finsp^H$ to compute $\yinsp \gets (H(\yeval^1,\cidi),
H(\yeval^1,\cidu))$. Thus, if the verifier $V$ wants to revoke a credential, all
it has to do is ask the opener to decrypt $\yinsp$, via $RetrieveCredHandle$.
%
In more detail, the algorithms are as follows:

\begin{description}
\item[$\parm \gets \Setup(1^\secpar)$.] Runs $\Setup(1^\secpar)$ as in
  \CUASGenInt.
\item[$(\ipk,\isk) \gets IssKeyGen(\parm)$.] Runs \KeyGen+\ISet from
  \CUASGenInt, setting $\fissue^1$ as issuance function.
\item[$(\ipk,\isk) \gets OKeyGen(\parm)$.] Runs \OKeyGen from \CUASGenInt,
  setting $\finsp^F$ as opening function.
\item[$(\upk,\usk) \gets UserKeyGen(\ipk)$.] Runs \KeyGen from \CUASGenInt.
\item[$\langle \cred/\bot,\top/\bot \rangle \gets
  \langle \Obtain(\usk,\ipk,\attrs),\Issue(\isk,\upk,\attrs) \rangle$.]
  Runs the $\langle \Obtain,\Issue \rangle$ protocol from \CUASGenInt, using
  $\fissue^1$ as issuance function.
\item[$\langle 1/0,(\yeval,\utrans)/(0,\utrans)\rangle \gets \langle Show(\ipk,
  \attrs,\dattrs,\cred),
  \Verify(\ipk,\dattrs) \rangle$.]
  Runs the interactive $\langle\Sign,\Verify\rangle$ of \CUASGenInt,
  using $\feval^{\dattrs,H,V,L}$ as signature evaluation function, and
  $\finsp^F$ as open function. The \yeval value output by the verifier, when
  it accepts, is the $\yeval^0$ value produced by \Sign in \CUASGenInt.
\item[$h/\bot \gets RetrieveCredHandle(\osk,\utrans)$.]
  Gets $(\sig,\yeval^0)$ from the $\langle Show,\Verify\rangle$ transcript
  \utrans, and runs \Open from \CUASGenInt. This returns $(H(V,\cidi),
  H(V,\cidu))$, which the verifier can add to its revocation list $L$.
\end{description}

\paragraph{Security of \CUASRAC.} %
We prove that \CUASRAC is an anonymous and unforgeable anonymous credential
scheme, according to \cite{fhs19}, if the underlying \CUASGen construction
has signature anonymity and is unforgeable.
%
As for \CUASAC, we reduce to security of \CUASGen rather than to
$\CUASGen^{int}$. The ``translation'' between \CUASGen and $\CUASGen^{int}$ is
again trivial.

\begin{theorem}[Anonymity of \CUASRAC]
  \label{thm:anon-cuasrac}
  If the base \CUASGen construction has signature anonymity according to
  \defref{def:sign-anonymity-uas}, then \CUASAC is an anonymous AC scheme
  according to \cite{fhs19}.
\end{theorem}

\begin{proof}
  Given \adv~ against anonymity of \CUASRAC as defined in
  \figref{fig:model-rac}, we build an adversary \advB against anonymity of
  \CUASGen as defined in \defref{def:sign-anonymity-uas}.
  %
  But, first observe that, since the revocation oracle is basically an extension
  of the \OPEN oracle -- also keeping track of the revoked users per verifier --
  queries involving challenge users are rejected if the revocation handle does
  not match for both challenge users.

  To simulate oracle calls by \adv, \advB simply redirects to its corresponding
  oracle, i.e.: for calls to $HU,CU,Obtain,Show,LoR,ReqCredHandle$, \advB
  redirects to \HUGEN, \CUGEN, \OBTAIN, \SIGN, \CHALb, or \OPEN (in this case,
  also keeping track of the queries, per verifier), respectively.
  Note that ``translating'' the inputs from the former into inputs for the
  latter is trivial. Also, observe that $ReqCredHandle$/\OPEN essentially
  refuses to open transcripts from challenge users (signatures, in the
  non-interactive equivalent), as $F(V,\ccidi_0) \neq F(V,\ccidi_1)$ (and,
  similarly, for $\ccidu_b$) with overwhelming probability (see the $\tyinsp
  \neq \yinsp$) check in \figref{fig:oracles2}. Obviously, if \adv~wins in its
  anonymity game, then so does \advB with the same probability.
  % 
  \qed
\end{proof}

\begin{theorem}[Unforgeability of \CUASRAC]
  If the base \CUASGen construction has unforgeability according to
  \defref{def:forge-uas}, and $F$ is one-way, then \CUASRAC is an
  unforgeable RAC scheme.
\end{theorem}

\begin{proof}
  Given \adv~against unforgeability of \CUASRAC as defined in
  \figref{fig:model-rac}, we build an adversary \advB against unforgeability of
  \CUASGen as defined in \defref{def:forge-uas}.

  To simulate oracle calls by \adv, \advB acts as in the anonymity proof. Assume
  that \adv~wins its game via the winning condition of line 9 in the RAC
  unforgeability game. The same reasoning as in \CUASAC applies here. Next, note
  that the probability of \adv~winning its game via condition $h_0 \in L$ at
  line 11 is negligible, due to soundness of the NIZK for $\NIZKRel_{\Sign}$.
  Moreover, soundness of NIZK for $\NIZKRel^{\Sign}$ as well as one-wayness
  of $F$ ensure that, if $h_1 \notin L$, then $b_1 = 1$. Namely, even when given
  $F(V,\cidi),F(V,\cidu)$ for a credential identified by $(\cidi,\cidu)$ owned
  by an honest user, and some (potentially malicious) $V$, it is not possible to
  generate $F(V',\cidi)$ nor $F(V',\cidu)$, for $V'=V$.
  % 
  \qed
\end{proof}

Note that, if we aim at global credential (resp. user) revocation rather than
verifier-local revocation, we can have \finsp just output the $(\cidi,\cidu)$
pair (resp., \upk), and no assumption regarding $F$ is needed.
